Rule-Based Normalization of German Twitter Messages

### 2013-08-13 darmstadt_gscl_2013/article_twitter_normalization.tex
misspelling section expanded

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{paralist} % for inparaenum
\usepackage{multirow} % for multirow
\usepackage{tabularx} % for centering in table
\usepackage{url}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\totaloov}{\% of total OOVs}
\newcommand{\uniqoov}{\% of unique OOVs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\pagestyle{headings}  % switches on printing of running heads
\title{MT @HanBaldwin: Fightin OOVs in German \#twitter}
\titlerunning{Makn Sens a Germa \#twitter}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Uladzimir Sidarenka \and Tatjana Scheffler \and Manfred Stede}
%
\authorrunning{Sidarenka et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Uladzimir Sidarenka, Tatjana Scheffler, Manfred Stede}
%
\institute{University of Potsdam,\\
  \email{{uladzimir.sidarenka@uni-potsdam.de,
      tatjana.scheffler@uni-potsdam.de,
      manfred.stede@uni-potsdam.de},\\ WWW home page:
    \texttt{http://www.ling.uni-potsdam.de/acl-lab/SocMedia/main.htm}}}

@@

We subdivided class \ref{dict} into the following subclasses:
\begin{enumerate}
\item regular German words, e.g. \textit{meins},
  \textit{losziehen};\label{regular}
\item compounds, e.g.  \textit{Altwein},
  \textit{Amtsapothekerin};\label{compound}
\item abbreviations, e.g. \textit{NBG}, \textit{OL};\label{abbr}
\item interjections, e.g.  \textit{aja}, \textit{haha};\label{inj}
\item named entities, with subclasses:\label{ne}
  \begin{enumerate}
  \item persons, e.g.  \textit{Ahmadinedschad}, \textit{Schweiger};
  \item geographic locations, e.g.  \textit{Biel}, \textit{Limmat};
  \item companies, e.g. \textit{Apple}, \textit{Facebook};
  \item product names, e.g. \textit{iPhone}, \textit{MacBook};
  \end{enumerate}
\item neologisms, with subclasses:\label{neolog}
  \begin{enumerate}
    \item newly coined German terms, e.g. \textit{entfolgen},
      \textit{gegoogelt};\label{new}
    \item loanwords, e.g. \textit{Community},
      \textit{Stream};\label{loan}
  \end{enumerate}
\item and, finally, foreign words like \textit{is} or \textit{now}
  which in contrast to \ref{loan} were not mentioned in any existing
  German lexica and did not comply with inflectional rules of German
  grammar.\label{fw}
\end{enumerate}
Though this division is admittedly arbitrary to a certain degree and
also has the disadvantage of simultaneously involving different
linguistic criteria, the underlying notion here was simple -- valid
words could have been omitted from an MRD either due to the
limitations of developers' capacities (group \ref{regular}), active
word formation processes or lexical productivity of the language
itself (groups \ref{compound} through \ref{new}) or also due to
language's openness to foreign language systems (groups \ref{loan} and
\ref{fw}).

@@

Luckily for us, this ``novelty'' by Messina brought along a strict
formal feature by which future hashtags could be identified. Other
types of words which are usually considered as OOVs in Twitter but
also have some unambiguous formal traits include at-tokens,
hyperlinks, e-mail addresses, and smileys. The presence of such formal
criteria suggests that these classes could best be handled by
rule-based methods and namely finite-state techniques like
finite-state transducers (FST) (cf. Jurafsky and Martin,
\cite{jurafsky}, pp.57-60).

For our purposes, we developed a prototypic Python system analogous to
an FST in which a set of regular expressions was associated with
corresponding actions performed on matched subgroups in cases when
these subgroups did not overlap. If some rules clashed together on
certain contexts, a set of special criteria was used to determine
which rule the preference should be given.

In this system, unambiguous tokens like web links and e-mail addresses
were replaced with an artificial word ``\%Mail''. Emoticons were
substituted by tokens ``\%PosSmiley'' or ``\%NegSmiley'', depending on
the type of emotion presumably conveyed by these expressions. In cases
when an emoticon was ambiguous with regard to its polarity, it was
replaced with the special word ``\%Smiley''. Furthermore, leading
``\#{}'' characters were stripped off from the beginning of hashtags
thus leaving only the alphabetic part of these tokens.

A more complicated case turned out to be the at-tokens and
hyperlinks. For them, we had to disambiguate whether these words
played a certain syntactic role in a sentence or could be easily
omitted without breaking the syntactic structure. In the former case,
these words were replaced with artificial counterparts. In the latter
case, such words were deleted from sentence in order not to confuse
further intended parsing processing.

Precision, recall and F1 score for replacement of Twitter-specific
phenomena is shown in Table \ref{table:tw-phenomen-metrics}.

@@

All the artificial words were added to custom user dictionaries of
\texttt{TreeTagger} and \texttt{hunspell}. For tagging, we assigned NE
tags to all artificial tokens with the only exception of emoticon
replacements which got the tag INJ. Furthermore, positions and lengths
of all replacements we remembered along with original words which were
replaced and a restoration step to original forms was performed for
certain classes of tokens after tagging.

@@

%
\section{Evaluation}\label{evaluation}

\section{Conclusion}\label{concl}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{paralist} % for inparaenum
\usepackage{multirow} % for multirow
\usepackage{tabularx} % for centering in table
\usepackage{url}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\totaloov}{\% of total OOVs}
\newcommand{\uniqoov}{\% of unique OOVs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\pagestyle{headings}  % switches on printing of running heads
\title{Rule-Based Normalization of German Twitter Messages}
\titlerunning{Rule-Based Normalization of German Twets}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Uladzimir Sidarenka \and Tatjana Scheffler \and Manfred Stede}
%
\authorrunning{Sidarenka et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Uladzimir Sidarenka, Tatjana Scheffler, Manfred Stede}
%
\institute{University of Potsdam,\\
  \email{{uladzimir.sidarenka@uni-potsdam.de,
      tatjana.scheffler@uni-potsdam.de,
      manfred.stede@uni-potsdam.de},\\ WWW home page:
    \texttt{http://www.ling.uni-potsdam.de/acl-lab/SocMedia/main.htm}}}

@@

We subdivided class \ref{dict} into the following subclasses:
\begin{enumerate}
\item regular German words, e.g. \textit{Piraterie},
  \textit{losziehen};\label{regular}
\item compounds, e.g.  \textit{Altwein},
  \textit{Amtsapothekerin};\label{compound}
\item abbreviations, e.g. \textit{NBG}, \textit{OL};\label{abbr}
\item interjections, e.g.  \textit{aja}, \textit{haha};\label{inj}
\item named entities, with subclasses:\label{ne}
  \begin{enumerate}
  \item persons, e.g.  \textit{Ahmadinedschad}, \textit{Schweiger};
  \item geographic locations, e.g.  \textit{Biel}, \textit{Limmat};
  \item companies, e.g. \textit{Apple}, \textit{Facebook};
  \item product names, e.g. \textit{iPhone}, \textit{MacBook};
  \end{enumerate}
\item neologisms, with subclasses:\label{neolog}
  \begin{enumerate}
    \item newly coined German terms, e.g. \textit{entfolgen},
      \textit{gegoogelt};\label{new}
    \item loanwords, e.g. \textit{Community},
      \textit{Stream};\label{loan}
  \end{enumerate}
\item and, finally, foreign words like \textit{is} or \textit{now}
  which in contrast to \ref{loan} were not mentioned in any existing
  German lexica and did not comply with inflectional rules of German
  grammar.\label{fw}
\end{enumerate}
Though this division is admittedly arbitrary to a certain degree and
also has the disadvantage of simultaneously involving different
linguistic criteria, the underlying notion here was simple -- valid
words could have been omitted from an MRD either due to the
limitations of developers' capacities (group \ref{regular}), active
word formation processes or lexical productivity of the language
itself (groups \ref{compound} through \ref{new}) or also due to
language's openness to foreign language systems (groups \ref{loan} and
\ref{fw}).

@@

Luckily for us, this Messina's ``novelty'' brought along a strict
formal feature by which future hashtags could be identified. Other
types of words which also are usually considered as OOVs in Twitter
but have unambiguous formal traits are at-tokens, hyperlinks, e-mail
addresses, and smileys. The presence of such formal criteria suggests
that these classes could best be handled by rule-based methods and
namely finite-state techniques like finite-state transducers (FST)
(cf. Jurafsky and Martin, \cite{jurafsky}, pp.57-60).

For our purposes, we developed a prototypic Python system analogous to
an FST in which a set of regular expressions was associated with
corresponding actions performed on matched subgroups. If some rules'
subgroups clashed together on certain contexts, a set of special
criteria was used to determine which rule had to be preferred. These
criteria involved length of text spans matched by subgroups, length of
left and right context mathced by corresponding regular expressions,
and, finally, the order in which these expressions appeared in rule
file.

In our system, unambiguous tokens like, for example, e-mail addresses
were replaced with an artificial word ``\%Mail''. Emoticons were
substituted by tokens ``\%PosSmiley'' or ``\%NegSmiley'', depending on
the type of emotion presumably conveyed by these expressions. In cases
when an emoticon was ambiguous with regard to its polarity, it was
replaced with the special word ``\%Smiley''. Furthermore, leading
``\#{}'' characters were stripped off from the beginning of hashtags
thus leaving only the alphabetic part of these tokens.

A more complicated case turned out to be the at-tokens and
hyperlinks. For them, we had to disambiguate whether these words
played an important syntactic role in sentence or not. In the former
case, these words were replaced with artificial counterparts. In the
latter case, such words were deleted from sentence in order not to
confuse further intended parsing processing.

Precision, recall and F1 metrics for replacement of Twitter-specific
phenomena are shown in Table \ref{table:tw-phenomen-metrics}.

@@

All the artificial words were added to custom user dictionaries of
\texttt{TreeTagger} and \texttt{hunspell}. For tagging, we assigned NE
tags to all artificial tokens with the only exception of emoticon
replacements which got the tag INJ. Furthermore, positions and lengths
of all replacements we remembered along with the original words which
were replaced, and a restoration step to original forms was performed
for user names after tagging.

@@

A closer look at unknown words which were classified as misspellings
revealed that a prevailing majority of them was considered as
colloquial spelling variants. Such spellings accounted for 70.5~\% of
detected spelling mistakes in \texttt{hunspell} data and 66.17~\% of
misspellings recognized by \texttt{TreeTagger}. Frequency distribution
of colloquial and non-colloquial misspellings for data analyzed with
either tools is shown in Figures \ref{fig:misspl-hunspell} and
\ref{fig:misspl-ttagger}.
...
%% \begin{minipage}[t]{0.45\textwidth}
%%   \begin{figure}[!ht]\label{fig:misspl-hunspell}
%%   \end{figure}
%% \end{minipage}

%% \begin{minipage}[t]{0.45\textwidth}
%%   \begin{figure}[!ht]\label{fig:misspl-ttagger}
%%   \end{figure}
%% \end{minipage}

As can be seen from the Figures, both kinds of spelling variations
have nearly Zipfian distribution. However colloquial spelling
variants, if used, tend to be used frequently whereas non-colloquial
misspellings are rather represented by sporadic singleton cases. This
could be explained by fact that colloquial spellings are usually
formed by systematic rewriting processes. These processes are supposed
to make written form of words more similar to their casual soundings
in everyday speech. According to our data, the most productive of such
processes were:
\begin{itemize}
  \item Omissions of `e' in unstressed (``schwa'') positions,
    e.g. \textit{w\"urd}, \textit{zuguckn} etc. In cases when `e' was
    part of the impersonal pronoun ``es'' and followed a verb,
    remaining `s' was usually appended to the preceding verb form;

  \item Complete omissions or replacement of final consonants with
    their voiceless equivalents, e.g. \textit{nich} instead of
    \textit{nicht} or \textit{Tach} instead of \textit{Tag};

  \item Omissions of `ei' from indefinite articles, e.g. \textit{ne}
    instead of \textit{eine} or \textit{nem} in lieu of
    \textit{einem};

  \item Multiple repetions of characters as a means of expressing
    elongation of sounds, e.g. \textit{Hilfeeee},
    \textit{s\"u\"u\"u\ss};
\end{itemize}

%
\section{Evaluation}\label{evaluation}
To evaluate our system, we used metrics

\section{Conclusion}\label{conclusion}
